### 2016-07-18 overview.tex
Starting with overlay fs
Final submission manuscript

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and
professors to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. Nevertheless, the
testbed description and procedures for setting it up are not restricted
to the academic scenario. All \ac{Raspi}s will be configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP}. This
eliminates the requirements of monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.

@@

In Fig.~\ref{fig:testbed_setup} and what follows, we will refer as the
testbed administrator to the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedure commands are
performed by the testbed administrator in a regular \ac{PC} running a
Linux distribution. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to later store it in a memory card. Thus, based on a
regular Raspbian Linux image, we explain how to resize and customize it to
include desired packages for performing routinary tasks in all the
\ac{Raspi}s. As another design consideration, when working on a testbed
with many devices, it may not always be desired to make persistent changes
on them. One scenario is when different users are interested in
running experiments on a rebooted testbed. Thus, we later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, an advanced step in our design considers to
fetch the \ac{Raspi}s root filesystem from a \ac{HTTP} server. This enables the
testbed administrator to put the data files in one location. The
\ac{Raspi}s will then get these data files after a reboot. We consider
that this step is particularly useful when there is a large amount
of devices to be configured and the setup process needs to be automated.
However, since a more complex scenario that might be required only in special
cases, we put this step in Section~\ref{sec:http} in the Appendix.

>>>>

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a 8 GB memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and
professors to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. Nevertheless, the
testbed description and procedures for setting it up are not restricted
to the academic scenario. All \ac{Raspi}s are configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP}. This
eliminates the requirements of monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.

@@

In Fig.~\ref{fig:testbed_setup} and what follows, we will refer as the
testbed administrator to the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedure commands are
performed by the testbed administrator in a regular \ac{PC} running a
Linux distribution. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to later store it in a memory card for each \ac{Raspi}.
Once configured, we store the resulting image file in a \ac{HTTP} server
as backup and in case the testbed administrator requires to make new
changes to this file. We also put all the required configuration files
and scripts for the \ac{Raspi}s setup in the \ac{HTTP} server so there is
a single place where system setup is stored and could be modified.
This simplifies the system maintenance at a later stage.
As another design consideration, when working on a testbed
with many devices, it may not always be desired to make persistent changes
on them. One scenario is when different users are interested in
running experiments on a rebooted testbed. Thus, we later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, we include a set of automation, monitoring
and cross-compilation tools over the top of our system in order to simplify
the execution of repetitive and long tasks, be able to follow the progress
of long task processes and compile relevant C++ source code for the testbed
administrator.

%an advanced step in our design considers to
% fetch the \ac{Raspi}s root filesystem from a \ac{HTTP} server. This enables the
% testbed administrator to put the data files in one location. The
% \ac{Raspi}s will then get these data files after a reboot. We consider
% that this step is particularly useful when there is a large amount
% of devices to be configured and the setup process needs to be automated.
% However, since a more complex scenario that might be required only in special
% cases, we put this step in Section~\ref{sec:http} in the Appendix.

### 2016-07-18 testbed_http.tex
Keep pushing
Testing procedure. BusyBox wget command fails when testing remote booting
Final submission manuscript

%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
<@\textcolor{gray}{
-------------------------------------------------------------}
\Reactivatenumber @>
<@\Suppressnumber @>dwc_otg.lpm_enable=0 console=serial0,115200 console=tty1 boot=http ip=dhcp
image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0 url=http://kom.aau.dk/project/TuneSCode/raspi/
rootfstype=ext4 elevator=deadline fsck.repair=yes rootwait quiet<@\Reactivatenumber @>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% ip=dhcp root=/dev/http http-options=hard,intr,ro

@@

\subsection{Initial RAM Filesystem for HTTP Fetch}
We enter again in \texttt{chroot} mode into the image to make the new
\texttt{initramfs} files. We note that for other testbed administrator
kernels, the files might be different. The commands for these tasks are:

@@

\subsection{Writing SD Cards for HTTP Fetch}

IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?

@@

Format partitions
\begin{lstlisting}[]
# mkfs.vfat -n BOOT ${HTTPDEV}p1
# mkfs.ext4 -L HOME ${HTTPDEV}p2
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\begin{lstlisting}[]
# umount --recursive ${HTTPROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

>>>>

%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
<@\textcolor{gray}{
-------------------------------------------------------------}
\Reactivatenumber @>
<@\Suppressnumber @>dwc_otg.lpm_enable=0 console=serial0,115200 console=tty1 boot=http ip=dhcp
image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0 url=http://kom.aau.dk/project/TuneSCode/raspi/
rootfstype=ext4 elevator=deadline fsck.repair=yes rootwait quiet<@\Reactivatenumber @>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%http://kom.aau.dk/project/TuneSCode/raspi/
% ip=dhcp root=/dev/http http-options=hard,intr,ro

@@

\subsection{Initial RAM Filesystem for HTTP Image Fetch}
We enter again in \texttt{chroot} mode into the image to make the new
\texttt{initramfs} files. We note that for other testbed administrator
kernels, the files might be different. The commands for these tasks are:

@@

\subsection{Writing SD Cards for HTTP Image Fetch}

%IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?
%ANSWER FROM NESTOR: YES! :)

We insert a test memory card and verify its associated device in the
filesystem to export it as a variable name:

@@

Format partitions. Before formating the partitions, take out the SD card
and plug it again!!!
\begin{lstlisting}[]
# mkfs.vfat -n BOOT ${HTTPDEV}p1
# mkfs.ext4 -L HOME ${HTTPDEV}p2
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\begin{lstlisting}[]
# umount ${HTTPROOTDIR}/boot
# umount ${HTTPROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

### 2016-07-18 tools.tex
Final submission manuscript

Then, the \texttt{\$CODEDIR/fabfile.py} file below provides a script with
some basic functionalities that can perform the few items above (i-iii).
In general, other administrators may require different functionalities
but that goes out of the scope of this work. Nevertheless, the following
file serves as a starting boilerplate:

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
<@\textcolor{gray}{
-------------------------------------}
\Reactivatenumber @>
from fabric.api import env, task, sudo
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

@@

To control a large set of devices, we simply need to include them in the
\texttt{env.hosts} list in the \texttt{\$CODEDIR/fabfile.py} file. Fabric has
many other functionalities that are useful in controlling a large set of
\ac{Raspi}s. For example, we may extract files or run automated
experiments. The included functionalities in the
\texttt{\$CODEDIR/fabfile.py} file will depend on the requirements of the
testbed administrator.

>>>>

Then, the \texttt{\$\{CODEDIR\}/fabfile.py} file below provides a script with
some basic functionalities that can perform the few items above (i-iii).
In general, other administrators may require different functionalities
but that goes out of the scope of this work. Nevertheless, the following
file serves as a starting boilerplate:

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{CODEDIR\}/fabfile.py}@>
<@\textcolor{gray}{
-----------------------------------------}
\Reactivatenumber @>
from fabric.api import env, task, sudo
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

@@

To control a large set of devices, we simply need to include them in the
\texttt{env.hosts} list in the \texttt{\$\{CODEDIR\}/fabfile.py} file.
Fabric has many other functionalities that are useful in controlling a
large set of \ac{Raspi}s. For example, we may extract files or run
automated experiments. The included functionalities in the
\texttt{\$\{CODEDIR\}/fabfile.py} file will depend on the requirements of the
testbed administrator.
